Bayesian multi-proxy reconstruction of latitudinal temperature gradients for the Early Eocene

### 2023-02-04 manuscript/manuscript.Rmd
Update manuscript.Rmd

# Introduction
The increasing availability of compilations and databases of deep-time climate data now offers the opportunity for empirical palaeoclimate reconstructions of unprecedented scope [e.g. @Veizer2015; @Hollis2019; @Grossman2022; @Judd2022]. In order for these reconstructions to be effective, two fundamental challenges need to be overcome: 1) disparate types of proxy data need to be integrated to inform a single climate reconstruction, and 2) large-scale scale climate patterns need to be inferred from localised, often sparse observations.

For the marine realm, most palaeoclimate reconstructions use geochemical proxy data

>>>>

# Introduction
The increasing availability of large compilations and databases of deep-time climate data now offers the opportunity for empirical palaeoclimate reconstructions of unprecedented scope [e.g. @Veizer2015; @Hollis2019; @Scotese2021; @Grossman2022; @Judd2022]. To exploit the full potential these data sets offer, two fundamental challenges need to be overcome: Disparate types of proxy data need to be integrated to inform a single climate reconstruction, and large-scale scale climate patterns need to be inferred from localised, often sparse observations. To address these problems, we present a novel Bayesian method that combines quantitative proxies and palaeoecological constraints into a fully quantitative model of the latitudinal temperature gradient, bridging spatial gaps in sparsely sampled climate data.

Existing palaeoclimate reconstructions are generally based on one of three approaches: 1) Quantitative, geochemical proxy data, 2) palaeobiological or geological climate indicators that provide only broad climate constraints which are hard to quantify, and 3) earth system modelling [@Burgener2023], with the latter ultimately relying on constraints provided by the geological and fossil record [@Lunt2021]. Combining quantitative proxy data with other climate constraints has the potential to improve upon climate reconstructions based on a single type of proxy [@Burgener2023], but is rarely employed in deep-time palaeoclimatic studies.

Climate data not based on geochemical, quantitative proxies generally provide broad categorical information, e.g. "tropical" or "dry", or broad quantitative constraints, e.g. "> 20 degC mean annual temperature". Integrating this information into fully quantitative climate reconstructions requires transforming it to a numerical scale. This approach is showcased by nearest living relative methods, commonly employed in terrestrial, Cenozoic climate reconstructions, which work on the premise that the climatic requirements of ancient plant taxa were similar to those of there nearest modern relatives [@Kvacek2007]. A plethora of methods have been developed to derive numerical climate values from taxonomic data [@Faucette2007; @...]. Approaches based on probability density functions have the advantage of naturally capturing the uncertainty associated with the climatic occupancy of modern nearest relatives [@Greenwood2017], and probability distributions can be readily integrated in a quantiative modelling framework.

Deep-time palaeoclimate reconstructions use geochemical proxy data
